---
title: "Class 6: R functions"
author: "Solomon Kim"
format: pdf
---

Functions are how we get work done in R. We call functions to do everything from reading data to doing analysis and outputing plots and results. 

All functions in R have at least 3 things: 

- a **name** (you get to pick this)
- input **arguments** (there can be onlyone or loads - again your call). 
- the **body** (where the work gets done, this code between teh curly pbrackets) 



## A first silly function

Let's write a function to add some numbers. we can call it `add()`

```{r}
x <- 10 
y <- 10 
x+y
```


```{r}
add <- function(x) { 
  y <- 10
  x + y
  }
```
 
 can I just use my new function?
 
```{r}
add(4)

```
 
 Let's make it a bit more flexible 
 
```{r}
add <- function(x,y) { 
  x + y
}

add(10,10)
```
```{r}
add <- function(x,y=10) { 
  x + y
}

add(10)
```
 
```{r}
add(10, 100)
```
 
 
 ## 2nd example grade() function
 
 Write a function to grade a student work. 
 
 We will start with a simple version of the problem and the following example student vector.
 
 
```{r}
# Example input vectors to start with
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)

sum(student1[1:8])/8
mean(student1)

```

 Start with student1
```{r}
mean(student1)
```
 
```{r}
mean(student2, na.rm=TRUE)
```
 
```{r}
mean(student3, na.rm=TRUE)
```

Ok lets try to work with student1 and find (and drop) the lowest score.

```{r}
student1
```

Google told me about min() and max()

```{r}
min(student1)
```

```{r}
which.min(student1)
```
```{r}
student1[which.min(student1)]
```


```{r}
student1[-8]
```


```{r}
mean(student1[-which.min(student1)])

```

Our first working snippet that drops the lowest score and calculates the mean

```{r}
mean(student1[-which.min(student1)])

```

Our approach to the NA probem (missing homeworks): We can replace all NA values with zero

1st task is find teh NA values( ex. where are they in the vectory)
```{r}
x <- student2
x
```
```{r}
x
is.na(x)
```

I have found the NA (TRUE) values from `is.na()` now I want to make them equal to zero (overwrite them/mask them etc.)

```{r}
x[is.na(x)] <-0
x
```
```{r}
y <- 1:5 
y
y[y>3] 
y[y>3] <-0
y
```



I want to combine the `is.na(x)` with making these elements equal to zero. And then take this "masked" (vector of student scores with NA values as zerio) and drop the lowest and get the mean

```{r}
x <- student2
x[is.na(x)] <-0
x

mean(x[-which.min(x)])

```


Now I can turn my most awesome snopet into my first function

```{r}

grade <- function(x) {
  # Make NA (missing work) equal fo zero
  x[is.na(x)] <-0
  # Drop lowest score and get mean 
  mean(x[-which.min(x)])
  
}
```

```{r}
grade(student3)
```


> Q1. Write a function grade() to determine an overall grade from a vector of student homework assignment scores dropping the lowest single score. If a student misses a homework (i.e. has an NA value) this can be used as a score to be potentially dropped. Your final function should be adquately explained with code comments and be able to work on an example class gradebook such as this one in CSV format: “https://tinyurl.com/gradeinput” [3pts]

```{r}
url <- "https://tinyurl.com/gradeinput"
gradebook <- read.csv(url, row.names =1)
head(gradebook)


```

The `apply()` function in R is super useful but can be a little confusing to begin with. Lets hae a look how it works. 

```{r}
ans <- apply(gradebook, 1, grade)
ans
```
>Q2. Using your grade() function and the supplied gradebook, Who is the top scoring student
overall in the gradebook? [3pts]


```{r}
which.max(ans)
max(ans)
```


>Q3. From your analysis of the gradebook, which homework was toughest on students (i.e. obtained the lowest scores overall? [2pts]

```{r}
which.min(apply(gradebook, 2, mean,  na.rm=TRUE))

```


>Q4. Optional Extension: From your analysis of the gradebook, which homework was most
predictive of overall score (i.e. highest correlation with average grade score)? [1pt]

```{r}
#ans
cor(gradebook$hw5, ans)
gradebook$hw5
```

Make all NA values into zero
```{r}
mask <- gradebook
mask[is.na(mask)] <- 0
mask
```

```{r}
cor(mask$hw5, ans)
```



Now we can use  `apply()` to examine the correlation of every assignmet in the masked gradebook to the overall score for each student in the class.

```{r}
apply(mask, 2, cor, y=ans)
```



```{r}
help(mean)
```



```{r}
r <- c(100, NA, 90, 90, 90, 90, 97, 80)
mean(na.omit(r))
mean(r, na.rm=TRUE)
```










